
Importing all packages and checking structure of the dataset. Checked class of variables, discarded invaluable columns and prepared and cleaned the data for further analysis.
```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library("tidyr")


data <- read_excel("C:/Users/SOMYA SINGHAL/Downloads/ANZ synthesised transaction dataset.xlsx")
head(data)

colnames(data)
nrow(data)

df<-data.frame(data)

#removed column 'card present flag', 'bpay biller code' and 'merchent code' as these contaied more than 30% of missing values
df<-select (df,-c(2,3,9))
df %>% View()

#understanding the data
summary(df)
str(df)

#checking unique values of the variables
unique(df$status)
unique(df$txn_description)
unique(df$movement)
unique(df$gender)
unique(df$age)
unique(df$merchant_state)
unique(df$currency)
class(df$date)
class(data$balance)

#factorise the columns
df$status<-factor(df$status,levels=c("authorized", "posted"))
df$txn_description<-factor(df$txn_description,levels=c("POS","SALES-POS","PAYMENT","INTER BANK","PAY/SALARY","PHONE BANK"))

#transform the state of date variable
df$date<-ymd(df$date)
class(df$date)

class(df$txn_description)
class(df$status)

```



Getting average transaction amount and no. of transactions which are made each month on average.
```{r}

mean(data$amount)

#extracted months from date column
month<-months(data$date)


df<-cbind(df,month)
head(df)
str(df)
colnames(df)

class(df$month)
unique(month)

aug <- subset(df,month=="August",transaction_id)
count(aug)
class(aug)
str(aug)

sep <- subset(df,month=="September",transaction_id)
count(sep)
class(sep)
str(sep)


oct <- subset(df,month=="October",transaction_id)
count(oct)
class(oct)
str(oct)


(count(aug)+count(sep)+count(oct))/3


#separating date and time 
df<- df %>% separate(extraction, into = c("transcation date", "transaction time"), sep = "T")
head(df)
colnames(df)
df$dates
df$time


```

## Including Plots

1) Plot for in what ways the transaction amount has been spend
2) Visualising location
3) Transaction amount spent monthly
4) Visualizing relation between age and amount
5) Visualizing relation between age and current balance

```{r}

#visualsing amount vs transaction method
boxplot(amount~txn_description,data=data, main="transaction vol vs spending",xlab="spending", ylab="transaction amount")

summary(df$txn_description)
summary(df$amount)

#removing the outliers
dframe<-df %>% group_by(txn_description) %>% filter(amount <= 80) 

#updated plot for amount vs transaction method
boxplot(amount~txn_description,data=dframe, main="transaction vol vs spending",xlab="spending", ylab="transaction amount",col=(c("red","green","pink","blue","orange")))


unique(df$merchant_state)

merch<-df %>% group_by(merchant_state) %>% filter(amount <= 70) 

boxplot(amount~merchant_state,data=merch,col=c("pink"))

date_value<-df %>% group_by(date) %>% count()

plot(date_value,df$amount,type="l",col='blue',ylab="amount",main="transaction amount per month")

plot(df$age,df$amount,type='l',col='red',xlab="age",ylab="amount",main="transaction amount vs age")

plot(df$age,df$balance,type='h',col='black',xlab="age",ylab="balance",main="balance vs age")

```


Calculating the annual salary for each customer

Correlation between salary and age
Correlation between salary and balance
Correlation between salary and gender

```{r}


library(tree)
library(ISLR)
library(rpart)
library(rpart.plot)
library(car)

salary<-df%>% filter(txn_description=="PAY/SALARY")%>% group_by(customer_id,age,gender,balance)%>% summarise(salary=mean(amount)*12)
head(salary)
class(salary)
typeof(salary)

cor(salary$salary,salary$age)
cor(salary$salary,salary$balance)

ggplot(salary,aes(x=age,y=salary))+geom_point(color= "steelblue",size=3)+ geom_smooth(method = "lm")

ggplot(salary,aes(x=gender,y=salary))+geom_point(color= "steelblue",size=3) 

boxplot(salary~gender,data=salary,main="corrolation between age and salary",col=c("pink","light blue"))

```


To check if gender influences salary income
```{r}


#dependency check btw salary n gender
leveneTest(salary ~ gender, data = salary)
t.test(salary~gender,data=salary,alternative = "greater", var.equal = FALSE)


salary$gender[salary$gender %in% "F"] <- 1
salary$gender[salary$gender %in% "M"] <- 0
salary$gender<-as.numeric(salary$gender)
typeof(salary$gender)


com=salary[,c("age","balance","salary","gender")]
cc = cor(com, method = "spearman")
cc


```


Building a simple linear model to predict annual salary based on age and gender
Building decision tree to predict annual salary and compare which gave better accuracy

```{r}
#regression model:

#Dividing data into training and testing

set.seed(100)  # setting seed to reproduce results of random sampling

# row indices for training data
trainingRowIndex <- sample(1:nrow(salary),0.8*nrow(salary))

 # model training data
trainingData <- salary[trainingRowIndex, ] 
testData  <- salary[-trainingRowIndex, ]

lmMod <- lm(salary ~ age+gender, data=salary)
summary(lmMod)

# build the model
distPred <- predict(lmMod, testData)

# make actuals_predicteds dataframe
actuals_preds <- data.frame(cbind(actuals=testData$salary, predicteds=distPred))  
correlation_accuracy <- cor(actuals_preds)
correlation_accuracy

head(actuals_preds)

#root mean square error
rmse <-print(sqrt(mean(distPred-testData$salary)^2))


########decision tree

sal<-salary[,c("age","salary","gender")]

#fitting the model
fit <- rpart(salary~., data = sal, method = 'anova')
rpart.plot(fit)

#predicting the based on test data
p<-predict(fit,testData,type="vector")

#root mean square value
rmse_tree<-sqrt(mean((testData$salary-p)^2))
rmse_tree

```

